\documentclass[10pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage{wrapfig}
\usepackage{beramono}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{fullpage}
\usepackage{cite}
\usepackage[top=1in, bottom=1in, left=0.85in, right=0.85in]{geometry}
\usepackage{bussproofs}
\usepackage{stmaryrd}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{url}
\usepackage{listings}
\usepackage{stmaryrd}
\usepackage{inconsolata}
\usepackage{xspace}
\usepackage{adjustbox}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{graphicx}

\usepackage{mathpartir}

\newcommand{\cL}{{\cal L}}
\newcommand{\sys}{\textsc{DopCert}\xspace}
\newcommand{\lbr}{\llbracket}
\newcommand{\rbr}{\rrbracket}
\newcommand{\gvd}{\Gamma \vdash}
\newcommand{\teq}{\triangleq}
\newcommand{\code}[1]{\texttt{\footnotesize #1}}



\newcommand{\flyingbox}[1]{\begin{flushleft}\fbox{{#1}}\end{flushleft}}
\newcommand{\concat}{\ensuremath{\!+\!\!\!\!+\!\,}}


\definecolor{forestgreen}{RGB}{64,139,64}
\lstdefinestyle{sql}{ % Define a style for your code snippet, multiple definitions can be made if, for example, you wish to insert multiple code snippets using different programming languages into one document
%backgroundcolor=\color{highlight}, % Set the background color for the snippet - useful for highlighting
language=sql,
basicstyle=\ttfamily, % The default font size and style of the code
breakatwhitespace=false, % If true, only allows line breaks at white space
breaklines=true, % Automatic line breaking (prevents code from protruding outside the box)
captionpos=b, % Sets the caption position: b for bottom; t for top
commentstyle=\color[rgb]{0,0.6,0}, % Style of comments within the code - dark green courier font
deletekeywords={VALUE, INPUT}, % If you want to delete any keywords from the current language separate them by commas
%escapeinside={\%}, % This allows you to escape to LaTeX using the character in the bracket
firstnumber=1, % Line numbers begin at line 1
frame=none, % Frame around the code box, value can be: none, leftline, topline, bottomline, lines, single, shadowbox
frameround=tttt, % Rounds the corners of the frame for the top left, top right, bottom left and bottom right positions
keywordstyle=\color{forestgreen},
morekeywords={PRODUCT}, % Add any functions no included by default here separated by commas
numbers=left, % Location of line numbers, can take the values of: none, left, right
numbersep=3pt, % Distance of line numbers from the code box
numberstyle=\large\color[rgb]{0.5,0.5,0.5}, % Style used for line numbers
rulecolor=\color{black}, % Frame border color
showstringspaces=false, % Don't put marks in string spaces
showtabs=false, % Display tabs in the code as lines
stepnumber=0, % The step distance between line numbers, i.e. how often will lines be numbered
tabsize=2, % Number of spaces per tab in the code
backgroundcolor=\color{white}
}
\lstset{escapeinside={@}{@}}

%for bussproofs abbreviation
\EnableBpAbbreviations

\newcommand\note[1]{\textcolor{red}{NOTE: #1}}

\begin{document}

\title{Bounded Verification of SQL Rewriting Rules}
\author{Chenglong Wang, Kaiyuan Zhang, Shumo Chu \\  
       Computer Science and Engineering\\ 
       University of Washington \\ 
       \{clwang, kaiyuanz, chushumo\}@cs.washington.edu  }
\date{}
\maketitle

\input{intro}
\input{overview}
\input{syntax}
\input{denotation}
\input{eval}

\section{Discussion}
\paragraph{Limitations} Though our technique is able to formally reason about the correctness of transformation rules, there are still two limitations that make our tool less practical.
\begin{itemize}
\item \emph{Scalability}: Our technique does not scale well enough to verify symbolic inputs with bigger size (currently the bounded symbolic input table is 5 different symbolic rows and 5 columns).
\item \emph{Table Schema Reasoning}: We didn't reason about table schema, and in practice, many useful rewriting rules are conditionally correct under certain schema constraints. And adding these features will make our tool more expressive to be able to capture more rewriting rules.
\end{itemize} 

\paragraph{Research Potential} Though our technique does not scale well with current SMT solver, there are still potential to make our tool practical with limited solver ability. And the key idea is to identify \emph{small model property}: if we cannot find a counter example with a symbolic input of with $m$ rows, we can prove that neither can we find a counter example with $n~(>m)$ symbolic rows, and thus we only need to verify the rewriting rule in a small symbolic input to ensure its correctness.

For example, Given the following two queries, if we can verify that given a $t$ with 1 symbolic rows, we can prove that there is no table $t$ with more than 1 rows can result in a counter example.
$$Q1: \code{SELECT t.a, t.b FROM t WHERE f1}$$
$$Q2: \code{SELECT t.a, t.b FROM t WHERE f2}$$

Thus, we believe that exploiting small model property on more complex SQL constructs will be the key to make our tool more practical.

\input{team}
\input{topic}

\bibliographystyle{plain}
\bibliography{writeup}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
